---
title: "Coral PCA and LDA"
author: "Ryan Botts, Elizabeth Maloney and Jason Chari"
date: "9/19/2019"
output: html_document
---

```{r setup, include = FALSE}
# had working directory issues (knitr doesn't agree with the terminal and none of the fixes seem to work)

```

```{r settings, include=FALSE}
ECHO_SETTING = FALSE #allows us to easily change setting in order to print all code, if necessary
INCLUDE_SETTING = FALSE #allows us to easily change setting in order to print all code, if necessary
EVAL_SETTING = FALSE #allos us to easily change setting to evalute all code, if necessary
CACHE_SETTING = FALSE

```

```{r packages, include=INCLUDE_SETTING, echo=ECHO_SETTING}
library(knitr)
library(stats)
#library(corrplot)
#library(plot3D)
#library(dendextend)
library(MASS)
library(plyr)
#library(wavelets)
library(cowplot)
library(RColorBrewer)
library(vegan) # used for PERMANOVA and computing dissimilarity matrix
library(ape)
library(ggfortify)
library(dplyr)
library(compositions) #as each row of areas sums to 1, we must consider this part of whole data or compositional data and thus we don't satisfy multivariate normality for MANOVA, we can overcome 
require(mvtnorm) # used to test for multivariate normality
require(PredPsych) # used for cross validation of LDA

library(DiscriMiner) # used for PLS-DA
library(caret)

# set base font size for graphics
theme_set(theme_grey(base_size = 18))
```

Read data from files.  There is a much better way to read these in, but there were many, many hiccups and we need to be done, so this gets it done.
```{r}
# there are differences between the woking directory for knitr and the project and strange things happened, so I am forcing the directory here, may not be necessary
#setwd("/Users/rbotts/Documents/Projects/Coral/CoralProject")
tempdir <- getwd()
pathtodat = paste0(tempdir,"/2016RawData/AllDCM/")

## @knitr data_in_dataframes
TRUNC_BEGIN = 30500#1 #first data point to use
TRUNC_END = 46500 #77001second data point to use

DELTA = 0.00033333333333 #Time increment by which intensity is measured (in minutes)

dfAllT_DCM <- read.csv("AllTCoralDat.csv")

dfAllT_DCM[,paste0("t",TRUNC_BEGIN:TRUNC_END)] <- sapply(dfAllT_DCM[,paste0("t",TRUNC_BEGIN:TRUNC_END)],as.numeric)
dfAllT_DCM[,1:6] <- sapply(dfAllT_DCM[,1:6],factor)

# create a short datasets of the categorical values for use in some of the functions that require X in one matrix and the classes in a data frame

df_DCM = data.frame(dfAllT_DCM[,c("Sample","Clade","Depth","Site", "SiteName")])#"SiteName"
df_DCM$Site = factor(df_DCM$Site, levels= 1:13, ordered = TRUE)
df_DCM$Clade = factor(df_DCM$Clade, levels = c("D","F","T"), labels = c("S. glaucum clade D","S. glaucum clade F","S. trocheliophorum"))
df_DCM$Depth <- as.numeric(df_DCM$Depth)
df_DCM$DepthCat <- NA
df_DCM[df_DCM$Depth<10,"DepthCat"] <- "0-10"
df_DCM[df_DCM$Depth>=10 & df_DCM$Depth<20,"DepthCat"] <- "10-20"
df_DCM[df_DCM$Depth>=20 ,"DepthCat"] <- "20-30"
df_DCM$DepthCat <- factor(df_DCM$DepthCat, levels = c("0-10","10-20","20-30"), ordered = TRUE)
df_DCM$SiteName <- as.factor(df_DCM$SiteName)
kable(table(df_DCM$SiteName, df_DCM$Clade))
kable(table(df_DCM$DepthCat, df_DCM$Clade))

```


Bin and normalize data from previous.  Hard coded from coralProject.R
```{r}
#computes the trapezoidal area of each time interval in the given vector using the delta as the height and the intensity readings of two successive rows as the side lengths
trap <- function(intens, delta) { 
  intens1 <- rep(intens)
  intens1 = intens1[-1]
  intens = intens[1:length(intens)-1]
  trapArea = (intens + intens1)*delta/2
  return(c(as.numeric(trapArea), 0))
}


#normalizes the vector input so that the total area adds to 1
norm <- function(trapArea){
  tot = sum(trapArea)
  return(trapArea/tot)
}

#returns a binned matrix whose bins are the sums of the areas df for each bin interval
#any remainder from ncol(df)/binwidth will not be included in the binned matrix
binner <- function(dfArea, dfAllT, binWidth, delta) {
  rowz = c()
  for(j in 1:nrow(dfArea)) {
    col = c()
    for(i in seq(1, ncol(dfArea)-binWidth, binWidth)){
      end = min(ncol(dfArea),i + binWidth - 1)
      col = c(col, sum(as.numeric(dfArea[j,i:end])))
    }
    rowz = rbind(rowz, col)
  }
  bins = 1:ncol(rowz)
  for (i in 1:ncol(rowz)){
    bins[i] = paste("Bin", as.character(i))
  }
  dimnames(rowz) = list(dfAllT[,1], bins)
  return(rowz)
}

# normalize each row requires the trapezoidal rule
dfDCM_area = dfAllT_DCM[,paste0("t",TRUNC_BEGIN:TRUNC_END)]
for(i in 1:nrow(dfDCM_area)){
  dfDCM_area[i, ] = norm(trap(as.numeric(dfDCM_area[i, ]), DELTA))
}
```




```{r chunkreading, include=INCLUDE_SETTING}


#function to standardize processing of data (trimming and binning)
procdata <- function(dfAllT, binwidth,  delta = DELTA){
  # dfAllT should be a matrix of only the chromatogram values
  df_area = dfAllT
  for(i in 1:nrow(df_area)){
    df_area[i, ] = norm(trap(as.numeric(df_area[i, ]), delta))
  }
  binned_matrix = binner(df_area, dfAllT, binwidth, delta)
  return(binned_matrix)
}

pcaWorkNew <- function(type, binned_matrix, grpdat){
 # binned matrix is a matrix of the trimmed and binned data
# grpdat is a data frame of categorigal variables for defining plots
  prc = prcomp(binned_matrix, center = TRUE, scale = FALSE)
  
  plot1 = autoplot(prc, data = grpdat, 
                   colour = 'Clade',  size = 4)+
    #ggtitle(paste0("PCA results for ", type, " with ",as.character(ncol(binned_matrix)), " bins")) +
    ggtitle(paste0(as.character(ncol(binned_matrix)+1), 
                   " bins")) +
    scale_color_manual(values=
                         c("#323efc","#fbb072","#6ac387"))+
    theme(
      axis.text.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      legend.text = element_text(face = "italic"),
      legend.position="none") 
  #+ #values = rev(colorRampPalette(brewer.pal(9, "BuGn"))(3)) +
                           
plot2 = autoplot(prc, data = grpdat, 
                 colour = 'Depth',  size = 4)+
    #ggtitle(paste0("PCA results for ", type, " with ",as.character(ncol(binned_matrix)), " bins"))  +
    ggtitle(" ") +
    #scale_color_brewer(palette="YlGn") +
    scale_color_gradient(low = "skyblue", 
                         high = "darkblue") + 
    theme(
      axis.text.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      legend.position="none")
  
  
  
plot3 = autoplot(prc, data = grpdat, 
                 colour = 'SiteName', size = 4) +
    #ggtitle(paste0("PCA results for ", type, " with ",as.character(ncol(binned_matrix)), " bins")) +
    ggtitle(" ") +
    scale_color_brewer(palette="Set1", name = "Site" ) + 
    theme(
      axis.text.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      legend.position="none")
  
  summary(prc)
  print(plot1)
  print(plot2)
  print(plot3)
  plot_grid(plot1,plot2,plot3,ncol = 3)
  return(plot_grid(plot1,plot2,plot3,ncol = 3))
}


pcoaPlots <- function(binned_matrix,grps){
    res <- pcoa(vegdist(binned_matrix, "bray", na.rm = TRUE))
    plot1 = ggplot(data = grps, 
                 aes(x=res$vectors[,1], 
                     y=res$vectors[,2],
                     color = Clade)) +
          geom_point() +
          ggtitle(paste0(as.character(ncol(binned_matrix)), " bins")) +
          scale_color_manual(values=c("#323efc","#fbb072","#6ac387")) +
          theme(axis.text.x = element_blank(),
              axis.text.y = element_blank(),
              axis.ticks = element_blank(),
              legend.text = element_text(face = "italic"),
              legend.position="none") +
          xlab("Axis 1") + ylab("Axis 2")

  temp1 <- plot1 + theme( legend.position="bottom") +
          guides(color=guide_legend(nrow=3,byrow=TRUE)) 
  
  # plot for PCoA by depth
 plot2 = ggplot(data = grps, 
                 aes(x=res$vectors[,1], 
                     y=res$vectors[,2], 
                     color = DepthCat)) +
              geom_point()  + 
              ggtitle(" ") +
              scale_color_manual(values=
                  scales::seq_gradient_pal("skyblue",
                        "darkblue", "Lab")(seq(0,1,
                        length.out=3))) +
              theme(axis.text.x = element_blank(),
                axis.text.y = element_blank(),
                axis.ticks = element_blank(),
                legend.position="none") +
              xlab("Axis 1") + ylab("Axis 2")
  
  temp2 <- plot2 + theme( legend.position="bottom") +
          guides(color=guide_legend(nrow=3,byrow=TRUE)) 
  
  plot3 = ggplot(data = grps, 
                 aes(x=res$vectors[,1], 
                     y=res$vectors[,2], 
                     color = SiteName)) + 
              geom_point()  + 
              ggtitle(" ") +
              scale_color_brewer(palette="Set1", name = "Site" ) + 
              theme(axis.text.x = element_blank(),
                axis.text.y = element_blank(),
                axis.ticks = element_blank(),
                legend.position="none") +
              xlab("Axis 1") + ylab("Axis 2")

  temp3 <- plot3 + theme( legend.position="bottom") +
          guides(color=guide_legend(nrow=3,byrow=TRUE)) 

  #plot_grid(plot1,plot2,plot3,ncol = 3)
  plot_grid(temp1,temp2,temp3,ncol = 3)
  return(plot_grid(plot1,plot2,plot3,ncol = 3))
}

ldaWorkNew <- function(binned_matrix, grp, predictStat = TRUE){
  # utility for systematically running the LDA, returns the LDA output for use in other functions
  # binned matrix refers to the matrix of binned chromatagram data
  # grp refers to the categorical response used in LDA
  # returns info from fitting to data
  lda_nocv = lda(grp ~ binned_matrix, CV=FALSE)
  lda_nocv_pred = predict(lda_nocv,binned_matrix)
  
  #lda_cv = lda(grp ~ binned_matrix, CV=TRUE)
  #lda_cv_pred = predict(lda_cv,binned_matrix)
  
  print(paste0("Train accuracy ",mean(lda_nocv_pred$class == factor(grp, ordered = FALSE))))
  if(predictStat){
    print("No crossvalidation LDA accuracy")
    print(confusionMatrix(lda_nocv_pred$class, grp))
  }  

  return(lda_nocv_pred) 
}

ldaPlots <- function(binned_matrix, grps, type = "DCM"){
  # utility for creating systematic LDA plots by clade, depth category and site
  # binned_matrix is a matrix of the processed chromat.
  # grps is a data frame of the categorical variables for the samples
  # plot for LDA output by clade
  ldaOutClade <- ldaWorkNew(binned_matrix, grps$Clade)
  plot1 = ggplot(data = grps, 
                 aes(x= ldaOutClade$x[,1],
                     y=ldaOutClade$x[,2],
                     color = Clade)) +
          geom_point() +
          stat_ellipse() +
          ggtitle(paste0(as.character(ncol(binned_matrix)+1), " bins")) +
          scale_color_manual(values=c("#323efc","#fbb072","#6ac387")) +
          theme(axis.text.x = element_blank(),
              axis.text.y = element_blank(),
              axis.ticks = element_blank(),
              legend.text = element_text(face = "italic"),
              legend.position="none") +
          xlab("LD1") + ylab("LD2")

  temp1 <- plot1 + theme( legend.position="bottom") +
          guides(color=guide_legend(nrow=3,byrow=TRUE)) 
  
  # plot for LDA by depth
  ldaOutDp <- ldaWorkNew(binned_matrix, grps$DepthCat)
  plot2 = ggplot(data = grps, 
                 aes(x= ldaOutDp$x[,1],y=ldaOutDp$x[,2], 
                     color = DepthCat)) +
              geom_point()  + 
              stat_ellipse() + 
              ggtitle(" ") +
              #scale_color_manual(values=
              #    c("#323efc","#fbb072","#6ac387")) +
    #scale_color_brewer(palette="YlGn", name = "Depth range") + 
              scale_color_manual(values=
                  scales::seq_gradient_pal("skyblue",
                        "darkblue", "Lab")(seq(0,1,
                        length.out=3))) +
              theme(axis.text.x = element_blank(),
                axis.text.y = element_blank(),
                axis.ticks = element_blank(),
                legend.position="none") +
              xlab("LD1") + ylab("LD2")
  
  temp2 <- plot2 + theme( legend.position="bottom") +
          guides(color=guide_legend(nrow=3,byrow=TRUE)) 
  
  ldaOutSite <- ldaWorkNew(binned_matrix, grps$SiteName)
  plot3 = ggplot(data = grps, 
                 aes(x= ldaOutSite$x[,1],y=ldaOutSite$x[,2], 
                     color = SiteName)) + 
              geom_point()  + 
              stat_ellipse() + 
              ggtitle(" ") +
              scale_color_brewer(palette="Set1", name = "Site" ) + 
              theme(axis.text.x = element_blank(),
                axis.text.y = element_blank(),
                axis.ticks = element_blank(),
                legend.position="none") +
              xlab("LD1") + ylab("LD2")

  temp3 <- plot3 + theme( legend.position="bottom") +
          guides(color=guide_legend(nrow=3,byrow=TRUE)) 
  # temporarily add labels for printable plots
  #temp1 = plot1 +
  #theme(legend.position="bottom") 
  #temp2 = plot2 +
  #theme(legend.position="bottom") 
  #temp3 = plot3 +
  #theme(legend.position="bottom") 
  #jpeg(paste0("LDAPlot",as.character(ncol(binned_matrix)),".jpg"), width = 1000, height = 300)
  plot_grid(plot1,plot2,plot3,ncol = 3)
  plot_grid(temp1,temp2,temp3,ncol = 3)
  #dev.off()
  
  return(plot_grid(plot1,plot2,plot3,ncol = 3))
}

plsDAPlots <- function(binned_matrix, grps, type = "DCM"){
  # utility for creating systematic LDA plots by clade, depth category and site
  # binned_matrix is a matrix of the processed chromat.
  # grps is a data frame of the categorical variables for the samples
  OutClade <- plsDA(binned_matrix, grps$Clade,autosel = FALSE,
                    comps = 3, cv = "LKO", k = 10)
  plot1 = ggplot(data = grps, 
                 aes(x= OutClade$components[,1],
                     y=OutClade$components[,2],
                     color = Clade)) +
          geom_point() +
          stat_ellipse() +
          ggtitle(paste0(as.character(ncol(binned_matrix)), " bins")) +
          scale_color_manual(values=c("#323efc","#fbb072","#6ac387")) +
          theme(axis.text.x = element_blank(),
              axis.text.y = element_blank(),
              axis.ticks = element_blank(),
              legend.text = element_text(face = "italic"),
              legend.position="none") +
          xlab("X 1") + ylab("X 2")

  temp1 <- plot1 + theme( legend.position="bottom") +
          guides(color=guide_legend(nrow=3,byrow=TRUE)) 
  
  # plot for LDA by depth
  OutDp <- plsDA(binned_matrix, grps$DepthCat,
                 autosel = FALSE, comps = 3, cv = "LKO", k = 10)
  plot2 = ggplot(data = grps, 
                 aes(x= OutDp$components[,1],
                     y=OutDp$components[,2], 
                     color = DepthCat)) +
              geom_point()  + 
              stat_ellipse() + 
              ggtitle(" ") +
              #scale_color_manual(values=
              #    c("#323efc","#fbb072","#6ac387")) +
    #scale_color_brewer(palette="YlGn", name = "Depth range") + 
              scale_color_manual(values=
                  scales::seq_gradient_pal("skyblue",
                        "darkblue", "Lab")(seq(0,1,
                        length.out=3))) +
              theme(axis.text.x = element_blank(),
                axis.text.y = element_blank(),
                axis.ticks = element_blank(),
                legend.position="none") +
              xlab("X 1") + ylab("X 2")
  
  temp2 <- plot2 + theme( legend.position="bottom") +
          guides(color=guide_legend(nrow=3,byrow=TRUE)) 
  
  OutSite <- plsDA(binned_matrix, grps$SiteName,
                   autosel = FALSE, comps = 5, cv = "LKO", k = 10)
  plot3 = ggplot(data = grps, 
                 aes(x= OutSite$components[,1],
                     y=OutSite$components[,2], 
                     color = SiteName)) + 
              geom_point()  + 
              stat_ellipse() + 
              ggtitle(" ") +
              scale_color_brewer(palette="Set1", name = "Site" ) + 
              theme(axis.text.x = element_blank(),
                axis.text.y = element_blank(),
                axis.ticks = element_blank(),
                legend.position="none") +
              xlab("X 1") + ylab("X 2")

  temp3 <- plot3 + theme( legend.position="bottom") +
          guides(color=guide_legend(nrow=3,byrow=TRUE)) 
  # temporarily add labels for printable plots
  #temp1 = plot1 +
  #theme(legend.position="bottom") 
  #temp2 = plot2 +
  #theme(legend.position="bottom") 
  #temp3 = plot3 +
  #theme(legend.position="bottom") 
  #jpeg(paste0("LDAPlot",as.character(ncol(binned_matrix)),".jpg"), width = 1000, height = 300)
  plot_grid(plot1,plot2,plot3,ncol = 3)
  plot_grid(temp1,temp2,temp3,ncol = 3)
  #dev.off()
  
  return(plot_grid(plot1,plot2,plot3,ncol = 3))
}

```

- Readings occur at delta = 0.00033333333333 seconds apart
- We align all samples so based on their first time point
- We omit the first 15000 time points (about 5 minutes) where there is little activity.
- We omit the last 15002 time points (about 5 minutes) where there is little activity.
- We take the trapezoidal area between two successive time point readings to approximate the area under the unknown curve between the given data points.
- We normalize the trapezoidal areas after truncating in order to appropriately compare the readings between samples.
- We bin the data to reduce the data's dimensionality. 
    + Each bin represents one explanatory varialbe.
    + We describe the bins by their "width" which is the number of time points included in each bin (or the total amount of time the bin's datapoints span)
- See DataNotes for additional details.

## Binning and preprocessing for analysis

As the initial binned data is normalized to 1 for each row, we must view the data as compositional, thus we use an Isometric Log Ratio transform so that we may perform MANOVA on it.  See here ![](https://link.springer.com/article/10.1023/A:1023818214614)

```{r}
raw1 = procdata(dfAllT_DCM[,paste0("t",TRUNC_BEGIN:TRUNC_END)],1600)
raw2 = procdata(dfAllT_DCM[,paste0("t",TRUNC_BEGIN:TRUNC_END)],1000)
raw3 = procdata(dfAllT_DCM[,paste0("t",TRUNC_BEGIN:TRUNC_END)],800)
raw4 = procdata(dfAllT_DCM[,paste0("t",TRUNC_BEGIN:TRUNC_END)],500)
raw5 = procdata(dfAllT_DCM[,paste0("t",TRUNC_BEGIN:TRUNC_END)],100)
rawAll = procdata(dfAllT_DCM[,paste0("t",TRUNC_BEGIN:TRUNC_END)],1)

# transform data due to initial data being compositional
d1 <- ilr(raw1)
d2 <- ilr(raw2)
d3 <- ilr(raw3)
d4 <- ilr(raw4)
d5 <- ilr(raw5)
```


# PCA analysis
## Analysis of all bin sizes
Create table of plots for all bin sizes
```{r}
bins1 = pcaWorkNew(type = "DCM", d1, df_DCM)
bins2 = pcaWorkNew(type = "DCM", d2, df_DCM) #1454
bins3 = pcaWorkNew(type = "DCM", d3, df_DCM)
bins4 = pcaWorkNew(type = "DCM", d4, df_DCM)
bins5 = pcaWorkNew(type = "DCM", d5, df_DCM)


#plot_grid(bins1, bins2, bins3, bins4, bins5, nrow = 5)

pdf("PCAPlots.pdf", width = 8, height = 14)
plot_grid(bins1, bins2, bins3, bins4, bins5, nrow = 5)
dev.off()
```
## Analyze the 32 bin case individually

Next we analyze the 32 bins only using PCA

We begin by performing MANOVA on the raw binned data to determine if the data peaks actually appear to be clear.  MANOVA verifies that Clade membership oes explain the variation, while depth category and site do not.  We verify that reducing the dimension does not change this.



```{r}

#Plot single plot for figure
prc = prcomp(d4, center = TRUE, scale = FALSE)
summary(prc)

# Select the principal components describing over 95% percent of the variance in the binned peaks.  For 32 bins, 8 principal components were selected.

summary(manova(cbind(d4)~Clade, data = df_DCM))
summary(manova(cbind(d4)~DepthCat, data = df_DCM))
summary(manova(cbind(d4)~SiteName, data = df_DCM))

# using MANOVA to assess how well the principal components are explained by each of the factors
summary(manova(cbind(prc$x[,1:8])~Clade, data = df_DCM))
summary(manova(cbind(prc$x[,1:8])~DepthCat, data = df_DCM))
summary(manova(cbind(prc$x[,1:8])~SiteName, data = df_DCM))

```
Generate the graphics showing the groupings by PCA on 32 bins
```{r}

plot1 = autoplot(prc, data = df_DCM, 
                 colour = 'Clade',  size = 4)+
    ggtitle(paste0("PCA results with ",
                   as.character(ncol(d4)+1), " bins")) +
    scale_color_manual(values=
                         c("#323efc","#fbb072","#6ac387"))+ 
    theme(
      axis.text.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      legend.text = element_text(face = "italic"),
      legend.position="bottom"
      ) + guides(colour=guide_legend(nrow=4,byrow=TRUE))
  
  plot2 = autoplot(prc, data = df_DCM, 
                   colour = 'Depth',  size = 4)+
    ggtitle(" ") +
    #scale_color_brewer(palette="YlGn", name = "Depth range") +
    scale_color_gradient(low = "skyblue", 
                         high = "darkblue") +
    theme(
      axis.text.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      legend.position="bottom")

  plot3 = autoplot(prc, data = df_DCM, 
                   colour = 'SiteName', size = 4) +
    ggtitle(" ") +
    scale_color_brewer(palette="Set1", name = 'Site') + 
    theme(
      axis.text.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      legend.position="bottom"
      ) + 
    guides(colour=guide_legend(nrow=4,byrow=TRUE))
  
 
  
  l1 = get_legend(plot1)
  l2 = get_legend(plot2)
  l3 = get_legend(plot3)
  plot1 = plot1 + theme(legend.position = "none")
  plot2 = plot2 + theme(legend.position = "none")
  plot3 = plot3 + theme(legend.position = "none")
  prow <- plot_grid(plot1,plot2,plot3, ncol = 3)
  lrow <- plot_grid(l1,l2,l3, ncol = 3)

pdf("PCAPlot32.pdf", width = 13, height = 5.5)
  #plot_grid(plot1,plot2,plot3,ncol = 3,rel_heights = c(1, .3))
plot_grid(prow,lrow, ncol = 1, rel_heights = c(1,.3))
  
dev.off()
  
```


# LDA analysis

## Analysis of all bin sizes
Generate table of plots
```{r}
ldaPlots(d1, df_DCM)
ldaPlots(d2, df_DCM)
ldaPlots(d3, df_DCM)
ldaPlots(d4, df_DCM)
ldaPlots(d5, df_DCM)
#p4 <- ldaPlots(d4, df_DCM)
pdf("LDAPlots.pdf", , width = 8, height = 14)
plot_grid(ldaPlots(d1, df_DCM),ldaPlots(d2, df_DCM),ldaPlots(d3, df_DCM),ldaPlots(d4, df_DCM),ldaPlots(d5, df_DCM), nrow = 5)
dev.off()
```

## Analysis of 32 bins only
LDA analysis for 32 bins 
```{r}
ldaOutClade <- ldaWorkNew(d4, df_DCM$Clade)
ldaOutDp <- ldaWorkNew(d4, df_DCM$DepthCat)
ldaOutSite <- ldaWorkNew(d4, df_DCM$SiteName)

# use caret confusion matrix statistics
#confusionMatrix(ldaOutClade$class, df_DCM$Clade)
#confusionMatrix(ldaOutDp$class, df_DCM$DepthCat)
#confusionMatrix(ldaOutSite$class, df_DCM$SiteName)

# cross validation analysis using the PredPsych package
LinearDA(data.frame(cbind(df_DCM$Clade, d4)), 1, cvFraction = .9, extendedResults = TRUE)
LinearDA(data.frame(cbind(df_DCM$DepthCat, d4)), 1, cvFraction = .9, extendedResults = TRUE)
LinearDA(data.frame(cbind(df_DCM$SiteName, d4)), 1, cvFraction = .9, extendedResults = TRUE)

```

figure for LDA for 32 bins
```{r}

plot1 = ggplot(data = df_DCM, 
                 aes(x=ldaOutClade$x[,1], 
                     y=ldaOutClade$x[,2],
                     color = Clade))+
          geom_point() +
          stat_ellipse() +
          ggtitle(paste0(
            as.character(ncol(d4)+1), " bins")) +
          scale_color_manual(values=
                  c("#323efc","#fbb072","#6ac387")) +
          xlab("LD1") + ylab("LD2") + 
          theme(axis.text.x = element_blank(),
                axis.text.y = element_blank(),
                axis.ticks = element_blank(),
                legend.text = element_text(
                  face = "italic"),
                legend.position="bottom"
              ) + 
          guides(color=guide_legend(nrow=4,byrow=TRUE))

  
# plot for LDA by depth

plot2 = ggplot(data = df_DCM, 
                 aes(x= ldaOutDp$x[,1],y=ldaOutDp$x[,2], 
                     color = DepthCat)) +
              geom_point()  + 
              stat_ellipse() + ggtitle(" ") +
              #scale_color_brewer(palette="YlBl", 
              #       name = 'Depth range (m)') +
              scale_color_manual(values=
                  scales::seq_gradient_pal("skyblue",
                        "darkblue", "Lab")(seq(0,1,
                        length.out=3))) + 
              xlab("LD1") + ylab("LD2") + 
              theme(axis.text.x = element_blank(),
                axis.text.y = element_blank(),
                axis.ticks = element_blank(),
              legend.position="bottom"
              ) + 
              guides(color=guide_legend(nrow=4,byrow=TRUE))
  

plot3 = ggplot(data = df_DCM, 
                 aes(x= ldaOutSite$x[,1],
                     y=ldaOutSite$x[,2], 
                     color = SiteName)) + 
              geom_point() +  
              stat_ellipse() + 
              ggtitle(" ") + 
              scale_color_brewer(palette="Set1", 
                                 name = 'Site')+
              xlab("LD1") + 
              ylab("LD2") + 
              theme(axis.text.x = element_blank(),
                axis.text.y = element_blank(),
                axis.ticks = element_blank(),
              legend.position="bottom"
              ) + 
              guides(color=guide_legend(nrow=4,byrow=TRUE))
l1 = get_legend(plot1)
l2 = get_legend(plot2)
l3 = get_legend(plot3)
plot1 = plot1 + theme(legend.position = "none")
plot2 = plot2 + theme(legend.position = "none")
plot3 = plot3 + theme(legend.position = "none")
prow <- plot_grid(plot1,plot2,plot3, ncol = 3)
lrow <- plot_grid(l1,l2,l3, ncol = 3)

pdf("LDAPlot32.pdf", width = 13, height = 5.5)
plot_grid(prow,lrow, ncol = 1, rel_heights = c(1,.3))
dev.off() 
#jpeg("LDAPlot32.jpg", width = 1000, height = 500)
#  plot_grid(plot1,plot2,plot3,ncol = 3)
#  dev.off()
```
# PCoA analysis
Alternatively it PCA use Principal Coordinates analysis.  Doesn't appear to show useful patterns.

## PCoA plots

```{r}
pcoaPlots(raw1, df_DCM)
pcoaPlots(raw2, df_DCM)
pcoaPlots(raw3, df_DCM)
pcoaPlots(raw4, df_DCM)
pcoaPlots(raw5, df_DCM)

pdf("PCoAPlots.pdf", width = 8, height = 14)
plot_grid(pcoaPlots(raw1, df_DCM),pcoaPlots(raw2, df_DCM),pcoaPlots(raw3, df_DCM),pcoaPlots(raw4, df_DCM),pcoaPlots(raw5, df_DCM), nrow = 5)
dev.off()
```


Compute the dissimilarity matrix using bray distance.  Run PERMANOVA to measure whether there are significant distances between groups.

```{r}
out = vegdist(raw4, "bray", na.rm = TRUE)
res <- pcoa(out)
summary(res)

adonis(out ~ df_DCM$Clade)
adonis(out ~ df_DCM$DepthCat)
adonis(out ~ df_DCM$SiteName)
```

```{r}

plot1 = ggplot(data = df_DCM, 
                 aes(x=res$vectors[,1], 
                     y=res$vectors[,2],
                     color = Clade))+
          geom_point() +
          ggtitle(paste0(
            as.character(ncol(raw4)), " bins")) +
          scale_color_manual(values=
                  c("#323efc","#fbb072","#6ac387")) +
          xlab("Axis 1") + ylab("Axis 2") + 
          theme(axis.text.x = element_blank(),
                axis.text.y = element_blank(),
                axis.ticks = element_blank(),
                legend.text = element_text(
                  face = "italic"),
                legend.position="bottom"
              ) + 
          guides(color=guide_legend(nrow=4,byrow=TRUE))

plot2 = ggplot(data = df_DCM, 
                 aes(x=res$vectors[,1], 
                     y=res$vectors[,2], 
                     color = DepthCat)) +
              geom_point()  + 
              ggtitle(" ") +
              #scale_color_brewer(palette="YlBl", 
              #       name = 'Depth range (m)') +
              scale_color_manual(values=
                  scales::seq_gradient_pal("skyblue",
                        "darkblue", "Lab")(seq(0,1,
                        length.out=3))) + 
              xlab("Axis 1") + ylab("Axis 2") + 
              theme(axis.text.x = element_blank(),
                axis.text.y = element_blank(),
                axis.ticks = element_blank(),
              legend.position="bottom"
              ) + 
              guides(color=guide_legend(nrow=4,byrow=TRUE))
  

plot3 = ggplot(data = df_DCM, 
                 aes(x=res$vectors[,1], 
                     y=res$vectors[,2], 
                     color = SiteName)) + 
              geom_point() +
              ggtitle(" ") + 
              scale_color_brewer(palette="Set1", 
                                 name = 'Site')+
              xlab("Axis 1") + 
              ylab("Axis 2") + 
              theme(axis.text.x = element_blank(),
                axis.text.y = element_blank(),
                axis.ticks = element_blank(),
              legend.position="bottom"
              ) + 
              guides(color=guide_legend(nrow=4,byrow=TRUE))
l1 = get_legend(plot1)
l2 = get_legend(plot2)
l3 = get_legend(plot3)
plot1 = plot1 + theme(legend.position = "none")
plot2 = plot2 + theme(legend.position = "none")
plot3 = plot3 + theme(legend.position = "none")
prow <- plot_grid(plot1,plot2,plot3, ncol = 3)
lrow <- plot_grid(l1,l2,l3, ncol = 3)

pdf("PCoAPlot32.pdf", width = 1000, height = 500)
plot_grid(prow,lrow, ncol = 1, rel_heights = c(1,.3))
dev.off() 

```



# PLSDA
```{r}
p1 <- plsDAPlots(raw1, df_DCM)
p2 <- plsDAPlots(raw2, df_DCM)
p3 <- plsDAPlots(raw3, df_DCM)
p4 <- plsDAPlots(raw4, df_DCM)
p5 <- plsDAPlots(raw5, df_DCM)
p6 <- plsDAPlots(rawAll, df_DCM)
#p4 <- ldaPlots(d4, df_DCM)
pdf("PLS-DAPlots.pdf", width = 8, height = 14)
plot_grid(p1,p2,p3,p4,p5,p6, nrow = 6)
dev.off()
```
```{r}
modAllClade <- plsDA(rawAll,df_DCM$Clade, autosel = FALSE, comps = 5, cv = "LKO", k = 10)
modAllClade$R2
modAllClade$Q2
modAllClade$error_rate

modAllDp <- plsDA(rawAll,df_DCM$DepthCat, autosel = FALSE, comps = 5, cv = "LKO", k = 10)
modAllDp$R2
modAllDp$Q2
modAllDp$error_rate

modAllSite <- plsDA(rawAll,df_DCM$SiteName, autosel = FALSE, comps = 5, cv = "LKO", k = 10)
modAllSite$R2
modAllSite$Q2
modAllSite$error_rate

```



```{r}

OutClade <- plsDA(raw4,df_DCM$Clade, autosel = FALSE, comps = 3, cv = "LKO", k = 10)
OutClade$R2
OutClade$Q2
OutClade$error_rate

OutDp <- plsDA(raw4,df_DCM$DepthCat, autosel = FALSE, comps = 3, cv = "LKO", k = 10)
OutDp$R2
OutDp$Q2
OutDp$error_rate

OutSite <- plsDA(raw4,df_DCM$SiteName, autosel = FALSE, comps = 3, cv = "LKO", k = 10)
OutSite$R2
OutSite$Q2
OutSite$error_rate
```
figure for PLS-DA for Full Data
```{r}

plot1 = ggplot(data = df_DCM, 
                 aes(x=modAllClade$components[,1], 
                     y=modAllClade$components[,2],
                     color = Clade))+
          geom_point() +
          stat_ellipse() +
          ggtitle(paste0(
            as.character("PLS-DA on unbinned data"))) + #ncol(raw4)), " bins")) +
          scale_color_manual(values=
                  c("#323efc","#fbb072","#6ac387")) +
          xlab("t1") + ylab("t2") + 
          theme(axis.text.x = element_blank(),
                axis.text.y = element_blank(),
                axis.ticks = element_blank(),
                legend.text = element_text(
                  face = "italic"),
                legend.position="bottom"
              ) + 
          guides(color=guide_legend(nrow=4,byrow=TRUE))

plot(plot1)
# plot for LDA by depth

plot2 = ggplot(data = df_DCM, 
                 aes(x= modAllDp$components[,1],
                     y=modAllDp$components[,2], 
                     color = DepthCat)) +
              geom_point()  + 
              stat_ellipse() + ggtitle(" ") +
              #scale_color_brewer(palette="YlBl", 
              #       name = 'Depth range (m)') +
              scale_color_manual(values=
                  scales::seq_gradient_pal("skyblue",
                        "darkblue", "Lab")(seq(0,1,
                        length.out=3))) + 
              xlab("t1") + ylab("t2") + 
              theme(axis.text.x = element_blank(),
                axis.text.y = element_blank(),
                axis.ticks = element_blank(),
              legend.position="bottom"
              ) + 
              guides(color=guide_legend(nrow=4,byrow=TRUE))
  

plot3 = ggplot(data = df_DCM, 
                 aes(x= modAllSite$components[,1],
                     y= modAllSite$components[,2], 
                     color = SiteName)) + 
              geom_point() +  
              stat_ellipse() + 
              ggtitle(" ") + 
              scale_color_brewer(palette="Set1", 
                                 name = 'Site')+
              xlab("t1") + 
              ylab("t2") + 
              theme(axis.text.x = element_blank(),
                axis.text.y = element_blank(),
                axis.ticks = element_blank(),
              legend.position="bottom"
              ) + 
              guides(color=guide_legend(nrow=4,byrow=TRUE))
l1 = get_legend(plot1)
l2 = get_legend(plot2)
l3 = get_legend(plot3)
plot1 = plot1 + theme(legend.position = "none")
plot2 = plot2 + theme(legend.position = "none")
plot3 = plot3 + theme(legend.position = "none")
prow <- plot_grid(plot1,plot2,plot3, ncol = 3)
lrow <- plot_grid(l1,l2,l3, ncol = 3)

pdf("PLS-DAPlotAll.pdf", width = 8, height = 3.5)
plot_grid(prow,lrow, ncol = 1, rel_heights = c(1,.3))
dev.off() 
```
